En Garde! Unguarded Iteration for Reversible Computation in the Delay Monad\thanks{Niccol\`o Veltri was supported by a research grant (13156) from VILLUM FONDEN.}

### 2019-07-01 mpc19/paper.tex
Minor change.

The uniform iteration operator introduces a notion of feedback,
typically used to model control flow operations such as while
loops. In the Kleisli category of the delay monad, this operation can
be intuitively described as follows: We can apply a function
$f : A \to B + A$ on an input $a$ and either produce an element
$b : B$, or produce a new element $a' : A$ which can be fed back to
$f$. This operation can be iterated, and it either terminates
returning a value in $B$ or it goes on forever without producing any
output. This form of iteration is ``unguarded'' because it allows the
possibility of divergence. The trace operator can then be seen as a
particular form of iteration where, given a function
$f : A + C \to B + C$, which can be decomposed as
$\mapL f : A \to B + C$ and $\mapR f : C \to B + C$, we first apply
$\mapL f$ on an input $a : A$, and, if the latter operation produces a
value $c : C$, we continue by iterating $\mapR f$ on $c$. Notice that
the notion of trace can be generally defined in monoidal categories
where the monoidal structure is not necessarily given by coproducts,
and it has been used to model other things besides iteration, such as
partial traces in vector spaces \cite{JoyalSV96}, though this use
falls outside of the scope of this paper.

>>>>

The uniform iteration operator introduces a notion of feedback,
typically used to model control flow operations such as while
loops. In the Kleisli category of the delay monad, this operation can
be intuitively described as follows: We can apply a function
$f : A \to B + A$ on an input $a : A$ and either produce an element
$b : B$, or produce a new element $a' : A$ which can be fed back to
$f$. This operation can be iterated, and it either terminates
returning a value in $B$ or it goes on forever without producing any
output. This form of iteration is ``unguarded'' because it allows the
possibility of divergence. The trace operator can then be seen as a
particular form of iteration where, given a function
$f : A + C \to B + C$, which can be decomposed as
$\mapL f : A \to B + C$ and $\mapR f : C \to B + C$, we first apply
$\mapL f$ on an input $a : A$, and, if the latter operation produces a
value $c : C$, we continue by iterating $\mapR f$ on $c$. Notice that
the notion of trace can be generally defined in monoidal categories
where the monoidal structure is not necessarily given by coproducts,
and it has been used to model other things besides iteration, such as
partial traces in vector spaces \cite{JoyalSV96}, though this use
falls outside of the scope of this paper.
